\documentclass[a4paper, 11pt, oneside]{book}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}   % Babel, replace with your language
\usepackage[T1]{fontenc}      % Font encoding
\usepackage{lmodern,cmap}     % lmodern font, correctly copyable characters in pdf

%\usepackage{hyperref}
\usepackage[hidelinks]{hyperref}
\usepackage{color}
\usepackage{setspace}
\usepackage{caption}

\usepackage{geometry}
\newgeometry{left=3.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}

\usepackage{amsmath}
\usepackage{amsfonts}

\makeatletter
\newif\if@restonecol
\makeatother

%ALGORITHM
\let\algorithm\relax
\let\endalgorithm\relax
\usepackage[linesnumbered,ruled,algonl,vlined]{algorithm2e}
\newcommand\mycommentfont[1]{\footnotesize \ttfamily \textcolor{blue}{#1}}
\SetCommentSty{mycommentfont}
\newcounter{algoline}

%GRAPHICS
\usepackage{graphicx}
\usepackage{epstopdf}
\DeclareGraphicsExtensions{.pdf,.eps,.png,.jpg,.mps}
\usepackage{placeins} % for \FloatbBarrier

%HEADER FOOTER
\usepackage{fancyhdr}
\fancyhf{}
\fancyhead[LO]{\slshape \rightmark}
\fancyhead[LE]{\slshape \leftmark}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0.0mm}
\renewcommand{\footrulewidth}{0.0mm}
\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{\markright{\thechapter.\ #1}{}} %\chaptername\ 
\renewcommand{\sectionmark}[1]{\markright{\thesection.\ #1}}
\renewcommand{\MakeUppercase}[1]{#1} % Get rid of CONTENTS BIBLIOGRAPHY upper cases


\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}

\frenchspacing \sloppy

\newcommand*{\perm}[2]{{}^{#1}\!P_{#2}}%
\newcommand*{\comb}[2]{{}^{#1}C_{#2}}%

\newcommand{\eat}[1]{}
\hyphenation{}

\begin{document}

\title{Ensemble Learning in Data Streams}


\author{
Hossain Mahmud \\
Technische Universit\"at M\"unchen\\
mahmud@in.tum.de\\
}

\date{}

%\maketitle
\pagenumbering{roman}
\onehalfspacing
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents
\clearpage

\listoftables
\addcontentsline{toc}{chapter}{List of Tables}
\clearpage

\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}
\clearpage

\listofalgorithms
\addcontentsline{toc}{chapter}{List of Algorithms}
\clearpage

%\normalsize
\onehalfspacing
\pagenumbering{arabic}
\setcounter{page}{1}


\input{introduction.tex}
\input{literature.tex}
\input{background.tex}

\chapter{Datasets}
Due to the lack of labeled of real-world data streams, stream mining algorithms are most of the time tested with synthesized data. Synthesized data has several advantages over poorly labeled real-world data streams. First of all, it is rather easy to produce. It has significantly less storage overhead. It is easier to control synthesizing parameters to generate data sets with desired concept drifting, recurring, etc. scenarios.

In this chapter, we briefly discuss some of such generators. For our experimentations we used random RBF generator. Thus, we also discuss the generation process and generated stream's properties of random RBF generator in greater details. We then describe the modifications in the generation process to introduce variable speed RBF streams.

\section{Stream Generators}
A number of generators have been introduced in literature in past decades: SEA concept generator, STAGGER concepts generator, rotating hyperplane generator, random RBF generator, waveform generator are some of the commonly used ones. 

\subsection{SEA Concepts Generator}
SEA concept generator is introduced in~\cite{street01:sea} to generate synthetic dataset with abrupt concept drift. It uses three parameters valued between $0$ and $10$ inclusive. The points of the dataset are divided into blocks of different concepts. Classification within the each block is controlled using using the input parameters along with a threshold value.

\subsection{STAGGER Concepts Generator}
STAGGER is one of the early methods developed for stream generation by Schlimmer et al.~\cite{schlimmer86:stagger}. Binary concepts are generated from three attributes. Attributes are size, shape, and color. Each of the attributes has three possible values. Small, medium, and large are the sizes; circle, triangle, and rectangle are the shapes; and red, blue, and green are the colors. $27$ combinations resulting from being mapped into a binary class by this generator.

\subsection{Rotating Hyperplane Generator}
Rotating Hyperplane generator is introduced in the evaluation of CVFDT~\cite{hulten01:cvfdt}. Hyperplanes can effectively be used to simulate concept drifting environments. A hyperplane in a $d$-dimensional space is a set of points $x$ that satisfies following equation:
\[
    \sum_{i=1}^{d} w_i x_i = w_0 = \sum_{i=1}^{d} w_i
\]
where $x_i$ is the $i$th coordinate of $x$. Hyperplane works as the boundary between two of the binary classes. By controlling the orientation and position of the hyperplane concept drift can be controlled. 

\subsection{Wavefront Generator}
Wavefront generator is a dataset available at UCI machine learning repository~\cite{internet:ucirepo}. Two or three base waves are used to generated a data set of three different classes of waveforms. The prediction task is to predict the classes of waveforms. The optimal Bayes classification rate is known to be $86\%$. There are two variations of this generator. Wave21 uses 21 numeric noisy attributes, and wave40 uses 19 more irrelevant attributes along with original 21.

\subsection{LED Generator}
LED generator is also available at UCI machine learning repository~\cite{internet:ucirepo}. The prediction task for this dataset is to predict the digit displayed on a seven segment LED display. Each segment has $10\%$ chance of being inverted. An optimal classification using naive Bayes on this dataset is found to $74\%$. 

\section{Random Radial Basis Function Generator}
All the generators mentioned above are relatively simple and easy to use. However, it is harder to model a complex scenario with those generators. Hypothesis space for these generator are not very large, except for rotating hyperplane generator. But with rotating hyperplane generator it is harder to overlapping space and outliers. Random radio basis function (RBF) generator is therefore introduced to generate complex concept type that is not straight forward to approximate, especially with decision tree model.



\section{Generation of Variable Speed RBF Stream}


\chapter{Algorithm}
\section{Problem Overview}
\section{Size Restricted Hoeffding Tree (SRHT)}
\section{Carry-Over Bagging with SRHT}

\chapter{Experimental Study}
\section{Test Study-case with Adult Dataset}
\section{Structural Study of Decision Trees}
\section{Performance Evaluation}
\subsection{Performance Metrics}
\subsection{Traditional Streams}
\subsection{Social Network-like Streams}

\chapter{Conclusion}
\section{Contribution}
\section{Future Works}
\section{Open Issues}

\bibliographystyle{apalike}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{ref}

\end{document}
