%\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}


\vspace*{2mm}

\epigraph{\textit{In the ecstasy of creation today\\
        Laughs my face, smile my eyes \\
        Glows my boiling blood \\
        In the brook of my shuttered soul \\
        The roaring tide brings the flood.} }{\textit{The Ecstasy of Creation \\ Kazi Nazrul Islam}}

%\epigraph{\textit{Persons attempting to find a motive in this narrative will be prosecuted;
%        persons attempting to find a moral in it will be banished;
%        persons attempting to find a plot in it will be shot.} }
%        {\textit{Adventures of Huckleberry Finn \\ Mark Twain}}


\vspace*{10mm}

\section* {Acknowledgments}
\vspace{7mm}
I would like to give my regards and thanks to all the persons who have become a part of this thesis.

\vspace{4mm} First of all, thanks to the Almighty for giving me the strength to finish this thesis.

\vspace{2mm} To my advisor, Dr. Eirini Ntoutsi, for her constant guidance, teachings, directions, corrections, supports, advices, instructions, inputs, and time to this thesis.

\vspace{2mm} To my advisor, Dr. Lothar Richter, for his guidance, directions, corrections, inputs, and time not just to this thesis, but also for teachings during Protein Prediction, and Data Mining courses.

\vspace{2mm} To my supervisor, Professor Dr. Burkhard Rost, for his insights into the thesis and for being an inspiration. Most importantly, for agreeing to be my supervisor when I wanted to do my thesis at LMU.

\vspace{2mm} To Professor Dr. Hans-Peter Kriegel, who is one of the reasons, I looked at TUM and LMU while choosing my graduate university. I am proud that I get to work in two of the world's finest research groups.

\vspace{2mm} To my former supervisor, Dr. M. Eunus Ali, for his guidance in choosing a thesis topic.

\vspace{2mm} To Christian Schulte zu Berge, for his support during my study at TUM.

\vspace{2mm} To my parents, bothers, sister-in-laws, and my nephews, for sending the love and support being a continent away.

\vspace{2mm} Lastly, to my friends here and abroad, especially Tanim, Ninoy, Charza, and Akram for, mostly, food and Euro trips.

